Reduced growth due to belowground sink limitation is not explained by reduced photosynthesis

### 2016-04-01 manuscript/manuscript_pve.Rmd
manuscript edits BM

\
Word counts
Main text:
Introduction:
Materials and methods:
Results:
Discussion:
Acknowledgements:
7 figures (6 color), 3 tables
Supporting information: 1 table

@@

\
(1)
$$P_i(t) = \frac{L_i(t-1)~Cday_i}~\sigma - R_i}{\epsilon}$$

where *L*~i~ is modelled total plant leaf area (m^2^), *C*~day,i~ is the predicted daily carbon assimilation (g C m^-2^ d^-1^, explained below), $\sigma$ is a self-shading parameter (explained below), $\epsilon$ is a biomass conversion efficiency parameter (g C g mass^-1^) and *R*~i~ is the mass based total respiration of all tissue components (g C d^-1^). Here, i represents model runs for each experimental treatment (1:7). Total respiration was calculated as

\
(2)
$$R_i(t) = \Sigma_{c=1}^{3}(R_\mathrm{c}~M_\mathrm{c,i})$$

where *R*~c~ is published or measured tissue respiration rates of each woody tissue component, including fine roots, coarse roots and stems (Table `r pvetabS`, g C g mass^-1^ d^-1^) and *M*~c,i~ is the modelled standing biomass of each component (g). Leaf respiration is included in the calculation of *C*~day~ (described below). The change in individual component biomass (*M*~c,i~), here solved on a daily time step, is given by

\
(3)
$$\frac{dM_\mathrm{c}}{dt} = \theta_\mathrm{c}~P_\mathrm{i}~-~(\Lambda_\mathrm{c}~M_\mathrm{c,i})$$
where $\theta$~c~ is the component specific biomass partitioning to whole plant biomass (%) and $\Lambda$~c~ is component specific turnover rate (yr^-1^). Because we did not observe branch turnover, $\Lambda$ of stems was assumed to equal 0. Total seedling biomass, per time step, was then equal to the sum of all biomass components; leaves, stems, fine roots and coarse roots.

\
*C*~day~ was predicted by using a coupled photosynthesis - stomatal conductance model [@farquhar1980biochemical;@medlyn2011reconciling] with the 'plantecophys' package [@duursma2015] in R with the mean photosynthetic parameters (*J*~max~, *V*~cmax~, *R* and *g*~1~) for each treatment and meteorological data from an onsite weather station. Values of *J*~max~ and *V*~cmax~ are mean estimates from AC~i~ curves over two measurement periods (explained above), *R* was empirically measured and the *g*~1~ parameter was generated by fitting the optimal stomatal conductance model from [@medlyn2011reconciling] to observed *g*~s~ values. Methods of the coupled leaf gas exchange model are also described in Duursma et al [-@duursma2014peaked]. Combined with the meteorological parameters; PPFD, air temperature, and relative humidity, at 15 min intervals, leaf *A*~n~ rates ($\mu$mol CO~2~ m^-2^ s^-1^) were then predicted for each soil volume treatment. *C*~day~ was calculated by converting predicted rates to mass C gain over 15 min time steps (g m^-2^) and then summed for 24 h. This resulted in 120 unique values of *C*~day~ for each soil volume treatment, one value for each day of the experiment. Thus, each daily time step for model runs included a value of *C*~day~ that represented both treatment specific photosynthetic parameters and meteorological constraints across the duration of the experiment.

\
It was further necessary to calculate a self-shading parameter ($\sigma$) when scaling leaf *A*~n~ with total plant leaf area. This was accomplished by utilizing 61 previously digitized *Eucalyptus* seedlings, covering 5 total species which include *E. tereticornis*, from Duursma et al. [-@duursma2012light] to run in 'YplantQMC' package [@YplantQMC] in R to build a 3D plant structure based on digitized metrics of plant allometry and crown structure. Using the same treatment specific physiological parameters listed above, 'YplantQMC' outputs total *A*~n~, using total leaf area, for seedlings assuming self-shading as well as for a full sun large horizontal leaf. The ratio of total *A*~n~ with self-shading to horizontal leaf was then used to calculate $\sigma$ for each of the 61 digitized seedlings, independently for each of the seven soil volume treatments. Next, the linear relationship between $\sigma$ and total leaf area was determined across digitized seedlings, within each treatment. For the growth model, $\sigma$ was then predicted for each daily time step using the previous days cumulative leaf area and this value  was then applied to *C*~day,i~. All default parameters used in model simulations are reported in Table `r pvetabS`.

\
We then used this model to test the hypothesis that the effects of belowground sink limitation on rates of leaf *A*~n~ were sufficient to accurately predict overall seedling biomass production after 120 days. First, a default model with mean photosynthetic parameters from the free seedlings was optimized to estimate a LMF which correctly predicted both the leaf mass and total biomass of the harvested free seedling (*M*~free~). This optimized LMF was then used to constrain model runs with treatment-specific *C*~day~, while keeping woody tissue respiration and tissue turnover parameters constant, to determine if changes in leaf *A*~n~ alone could predict biomass of seedlings in containers (*M*~pots~). Cumulative net leaf C gain for each treatment was equal to the sum of each value of *C*~day,i~ over 120 days and final seedling C was assumed to equal half of the final dry mass for both modeled and observed seedlings. Seedling biomass production was compared between model output and harvested seedlings with treatment-specific mean *C*~day~ by first scaling values to the free seedling control.

@@

# Acknowledgements
We thank Burhan Amiji for his outstanding technical assistance. We would like to thank the many Hawkesbury Institute for the Environment staff and students who helped during the experimental harvest.

>>>>

\
Word counts
\
Main text:
\
Introduction:
\
Materials and methods:
\
Results:
\
Discussion:
\
Acknowledgements:
\
7 figures (6 color), 3 tables
\
Supporting information: 1 table

@@

\
(1)
$$P_i(t) = \frac{L_\mathrm{i}~Cday_\mathrm{i}~\sigma - R_\mathrm{i}}{\epsilon}$$

where *L*~i~ is modelled total plant leaf area (m^2^), *Cday*~i~ is the predicted daily carbon assimilation (g C m^-2^ d^-1^, explained below), $\sigma$ is a self-shading parameter (explained below), $\epsilon$ is a biomass conversion efficiency parameter (g C g mass^-1^) and *R*~i~ is the mass based total respiration of all tissue components (g C d^-1^). Here, i represents model runs for each experimental treatment (1:7). Total respiration was calculated as

\
(2)
$$R_i(t) = \Sigma_{c=1}^{3}(R_\mathrm{c}~M_\mathrm{c,i}(t))$$

where *R*~c~ is published or measured tissue respiration rates of different woody tissue components (Table `r pvetabS`, g C g mass^-1^ d^-1^) and *M*~c,i~ is the modelled standing biomass for each component (g). Here, c represents fine roots, coarse roots or stems. Leaf respiration is included in the calculation of *Cday* (described below). The change in individual component biomass (*M*~c,i~), here solved on a daily time step, is given by

\
(3)
$$\frac{dM_\mathrm{c}}{dt} = \theta_\mathrm{c}~P_\mathrm{i}(t)~-~(\Lambda_\mathrm{c}~M_\mathrm{c,i})$$

where $\theta$~c~ is the component specific biomass partitioning to whole plant biomass (%) and $\Lambda$~c~ is component specific turnover rate (yr^-1^). Because we did not observe branch turnover, $\Lambda$ of stems was assumed to equal 0. Total seedling biomass, per time step, was then equal to the sum of all biomass components; leaves, stems, fine roots and coarse roots.

\
*Cday* was predicted by using a coupled photosynthesis - stomatal conductance model [@farquhar1980biochemical;@medlyn2011reconciling] with the 'plantecophys' package [@duursma2015] in R with the mean photosynthetic parameters (*J*~max~, *V*~cmax~, *R* and *g*~1~) for each treatment and meteorological data from an onsite weather station. Values of *J*~max~ and *V*~cmax~ are mean estimates from AC~i~ curves over two measurement periods (explained above), *R* was empirically measured and the *g*~1~ parameter was generated by fitting the optimal stomatal conductance model from [@medlyn2011reconciling] to observed *g*~s~ values. Methods of the coupled leaf gas exchange model are also described in Duursma et al [-@duursma2014peaked]. Combined with the meteorological parameters; PPFD, air temperature, and relative humidity, at 15 min intervals, leaf *A*~n~ rates ($\mu$mol CO~2~ m^-2^ s^-1^) were then predicted for each soil volume treatment. *Cday* was calculated by converting predicted rates to mass C gain over 15 min time steps (g m^-2^) and then summed for 24 h. This resulted in 120 unique values of *Cday* for each soil volume treatment, one value for each day of the experiment. Thus, each daily time step for model runs included a value of *Cday* that represented both treatment specific photosynthetic parameters and meteorological constraints across the duration of the experiment.

\
It was further necessary to calculate a self-shading parameter ($\sigma$) when scaling leaf *A*~n~ with total plant leaf area. This was accomplished by utilizing 61 previously digitized *Eucalyptus* seedlings, covering 5 total species which include *E. tereticornis*, from Duursma et al. [-@duursma2012light] to run in 'YplantQMC' package [@YplantQMC] in R to build a 3D plant structure based on digitized metrics of plant allometry and crown structure. Using the same treatment specific physiological parameters listed above, 'YplantQMC' outputs total *A*~n~, using total leaf area, for seedlings assuming self-shading as well as for a full sun large horizontal leaf. The ratio of total *A*~n~ with self-shading to horizontal leaf was then used to calculate $\sigma$ for each of the 61 digitized seedlings, independently for each of the seven soil volume treatments. Next, the linear relationship between $\sigma$ and total leaf area was determined across digitized seedlings, within each treatment. For the growth model, $\sigma$ was then predicted for each daily time step using the previous days cumulative leaf area and this value  was then applied to *Cday*~i~. All default parameters used in model simulations are reported in Table `r pvetabS`.

\
We then used this model to test the hypothesis that the effects of belowground sink limitation on rates of leaf *A*~n~ were sufficient to accurately predict overall seedling biomass production after 120 days. First, a default model with mean photosynthetic parameters from the free seedlings was optimized to estimate a leaf mass fraction (LMF) which correctly predicted both the leaf mass and total biomass of the harvested free seedling (*M*~free~). This optimized LMF was then used to constrain model runs with treatment-specific *Cday*, while keeping woody tissue respiration and tissue turnover parameters constant, to determine if changes in leaf *A*~n~ alone could predict biomass of seedlings in containers (*M*~pots~). Cumulative net leaf C gain for each treatment was equal to the sum of each value of *Cday*~i~ over 120 days and final seedling C was assumed to equal half of the final dry mass for both modeled and observed seedlings. Seedling biomass production was compared between model output and harvested seedlings with treatment-specific mean *Cday* by first scaling values to the free seedling control.

@@

# Acknowledgements
We thank Burhan Amiji for his outstanding technical assistance. We would like to thank the many Hawkesbury Institute for the Environment staff and students who helped during the experimental harvest. This research was funded by a PhD Scholarship drom the Hawkesbury Institute for the Environment (Western Sydney University).
